---
title: "Assignment Week 1"
subtitle: "Minor Applied Data Science: Association rule analysis"
output: html_document
---


The output to be handed in should be in `html`.

The data set in the file `Online Retail BIM.dat` consists of the invoice number
(first column) and the binary incidence matrix of items that have been sold 
in the transactions occurring between *01/12/2010* and *09/12/2011* for a 
UK-based and registered non-store online retail. The company mainly sells 
unique all-occasion gifts. Many customers of the company are wholesalers.
The number of transactions is 20511 and the number of disctinct items sold
is *4175*. Item information is given in the Word document
`Online Retail Items.docx`.

Use the R program to perform a Market Basket analysis. Save all your files
in one folder and use R script to save your work. Import the data by executing
`DAT<-read.table(‘Online Retail BIM.dat’,header=T)`

```{r, message=FALSE}
DAT <- read.table("Online Retail BIM.dat", header = T)
```

Load the packages `arules` and `arulesViz`. To be able to analyze the binary 
incidence matrix using the function `apriori()` it must be of 
class `transactions`. 

```{r, message=FALSE}
library(arules)
library(arulesViz)
```

First, the binary incidence matrix can be obtained from data frame DAT by 
removing the first column, which contains the invoice number, and then 
converting it to a matrix. Save the resulting matrix with name BIM by executing
`BIM<-as.matrix(DAT[-1])`

```{r}
BIM <- as.matrix(DAT[-1])
```

Next, the matrix is made of class transactions by executing
`BIM<-as(BIM,‘transactions’)`

```{r}
BIM <- as(BIM, "transactions")
```

BIM can now be used to find association rules.

**1. Plot the relative frequencies (the support) of the items and determine 
which three items have been sold most often.**

```{r}
itemFrequencyPlot(BIM,topN=3,type="relative", main="Rel. Item Freq. Plot")

# WHITE HANGING HEART T-LIGHT HOLDER[i3978], JUMBO BAG RED RETROSPOT [i1919] 
# and REGENCY CAKESTAND 3 TIER [i2952] have the top 3 highest relative frequency
```

**2. Determine which two association rules containing only two items have the 
highest support. Be aware of the defaults of the function apriori(). These 
defaults can be checked by executing ?apriori and reading the text for the 
parameter argument.**

```{r, warning=FALSE}
rules_2_items <- apriori(BIM, parameter = list(supp = .01,
                                       conf = .5,
                                       minlen = 2,
                                      maxlen = 2))

inspect(sort(rules_2_items, decreasing = T, by = "support")[1:5])

# items {i1917, i1919} have the highest support of 0.0402. 
# Their names are {JUMBO BAG PINK POLKADOT, JUMBO BAG RED RETROSPOT}

```

**3. How many items does the association rule with the highest confidence 
contain?**

```{r}
highest_confidence <- apriori(BIM, parameter = list(supp = .01,
                                       conf = .5,
                                       minlen = 2,
                                      maxlen = 10))




inspect(sort(highest_confidence, decreasing = T, by = "confidence")[1:5])

# Association with the highest confidence of 0.9587 contains 3 items which
# are antecedents (i1884, i3617) and consequent i1151 

# Antecedent names are {JAM MAKING SET PRINTED, SUKI  SHOULDER BAG} and the
# Consequent name is DOTCOM POSTAGE
```

**4. Which items are important antecedents of item 4050? What does this mean?**

```{r}
antecedent_4050 <- apriori(BIM, parameter = list(supp = .01,
                                                 conf = .5, 
                                                 minlen = 2),
                           appearance = list(default = "lhs",rhs="i4050"))

inspect(head(sort(antecedent_4050, decreasing = T, by = "confidence")))
# Antecedents are the items that appear on the left-hand side of an 
# association rule. They are also called premise or the condition of the rule 
# and are used to predict the consequent

# items i4042 and i4049 are important antecedents of i4050 because they have 
# the highest confidence.

# A purchase of antecedents WOODEN HEART CHRISTMAS SCANDINAVIAN and WOODEN STAR 
# CHRISTMAS SCANDINAVIAN most likely leads to a purchase of consequent 
# WOODEN TREE CHRISTMAS SCANDINAVIAN
```


**5. Which items are important consequents of item 1655? What does this mean?**

```{r}
consequent_1655 <- apriori(BIM, parameter = list(supp = 0.01,
                                                 conf = 0.1),
                           appearance = list(lhs = "i1655", default = "rhs"))

inspect(head(sort(consequent_1655, decreasing = T, by = "confidence")))
# Consequent are the items that appear on the right-hand side of an 
# association rule. They are also called the conclusion or the outcome of 
# the rule. They are predicted by the antecedent.

#A purchase of item i1655 most likely leads to a purchase of all these items 
#in the rhs specifically item i3051 which has the highest confidence of 0.7566

# A purchase of antecedent GREEN REGENCY TEACUP AND SAUCER most likely leads 
# to a purchase of consequent ROSES REGENCY TEACUP AND SAUCER
```

**6. Are there two items that are substitute to each other? Why?**

```{r}
item_substitutes <- apriori(BIM, parameter = list(supp = 0.01,
                                             conf = 0.5))

inspect(head(sort(item_substitutes, decreasing = F, by = 'lift')))

#No items are substitutes for each other as no combinations have a 
# lift less than 1
```

**7. Make a scatter plot and a two-key plot for association rules with minimum 
support 0.01 and minimum confidence 0.5.**

```{r, message=FALSE}
rules_plot <- apriori(BIM, parameter = list(support = 0.01, 
                                            confidence = 0.5))

plot(rules_plot, measure = c("support", "confidence"), shading = "lift")

plot(rules_plot, method = "two-key plot")
```

